### 2013-02-12 doc/icml2013/algo.tex
minor updates

@@

\subsection{Notation}

We have already defined $y$ to be the response variable and $x \in
\Re^d$ to be the covariates it depends on. Note that the matrix of
regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
$\beta_k$.

We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use $\odot$ to denote
the general dot product between $p$-th order tensors, i.e. $X \odot
Y = \sum_{i_1, i_2, \ldots i_p} X_{i_1, i_2, \ldots i_p} Y_{i_1, i_2,
\ldots i_p}.$ Similar to how $x^T A$ is the projection of a vector $x$
onto the row space of a matrix $A$, when $X \in \Re^{d^p}$ and $Y \in
\Re^{d^{p+q}}$,  with $p, q > 0$, $X \odot Y \in \Re^{d^q}$ is the
projection of $X$ onto the first $p$ dimensions of $Y$, i.e. $(X \odot
Y)_{i_1, \ldots, i_q} = \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p}
Y_{j_1, \ldots, j_p, i_1, \ldots, i_q}$.

The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

\subsection{Identifiability}

\subsection{Overview}

@@

To work around this problem, we will use regression on the conditional
moments, $\E[y | x]$ to recover the expected moments of the regressors,
$M_2$ and $M_3$. Recovering the $\beta_k$ and $\pi_k$ given these
symmetric moments has been addressed in prior work
\citep{AnandkumarHsuKakade2012, AnandkumarGeHsu2012}.  This algorithm as
been summarized in \algorithmref{algo:spectral-experts}.

@@

Let $\epsilon \sim \normal{0, \sigma^2}$ be independent Gaussian noise.
we have,
\begin{eqnarray*}
  y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
  \E[ y | x ] &=& M_1^T x \\
  \E[ y^2 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^2 + \E[ \epsilon^2 ] \\
  &=& M_2 \odot x^{\otimes 2} + \sigma^2 \\
  \E[ y^3 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^3
            + 3 \E[\epsilon^2] (\sum_{k=1}^{K} \pi_{k} \beta_k^T) x \\
  &=& M_3 \odot x^{\otimes 3} + 3 \sigma^2 M_1.
\end{eqnarray*}

% Details about low rank regression.
The symmetric structure of $M_2$ and $M_3$ implies that there are
$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
respectively. However, the moments are further constrained to have
a rank of $K$. To exploit this property, we will use trace norm
regularization with a proximal subgradient descent
algorithm\citationneeded to recover $M_2$,
\begin{eqnarray*}
  M_2 &=& \arg\min_{M} \half \sum_{i=1}^N ((\xn{i})\tp{2} \odot M - (\yn{i})^2)^2 + \lambda \|M\|_*.
\end{eqnarray*}

@@

\subsubsection{Implications for Identifiability}

@@

\subsection{Recovering $B$ and $\pi$}

Having recovered the moments $M_2$ and $M_3$, we must now solve for the
actual regressors $\beta_k$ and mixture weights $\pi_k$, which are the
eigenvectors and eigenvalues of the symmetric moments respectively,
\begin{eqnarray*}
  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
\end{eqnarray*}

>>>>

The goal is to recover the parameters $(\pi, B)$.
What should the moments be?

OUTLINE

\subsection{Recovering $B$ and $\pi$ from $M_2,M_3$}

If we knew the following quantities, then we could apply tensor factorization
\cite{AnandkumarGeHsu2012}.

Having recovered the compound parameters $M_2$ and $M_3$, we must now solve for the
actual parameters $\{\beta_k\}$ and $\{\pi_k\}$, which are the
eigenvectors and eigenvalues of the symmetric moments respectively,
\begin{eqnarray*}
  M_2 &=&  \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 2}\\
  M_3 &=& \sum_{k=1}^{K} \pi_k \beta_k^{\otimes 3}.
\end{eqnarray*}

Tensor power method

@@

We want to estimate from low-order moments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

To work around this problem, we will use regression on the conditional
moments, $\E[y \mid x]$ to recover the expected moments of the regressors,
$M_2$ and $M_3$. Recovering the $\beta_k$ and $\pi_k$ given these
symmetric moments has been addressed in prior work
\citep{AnandkumarHsuKakade2012, AnandkumarGeHsu2012}.  This algorithm as
been summarized in \algorithmref{algo:spectral-experts}.

@@

Let $\epsilon \sim \normal{0}{\sigma^2}$ be independent Gaussian noise.
we have,
\begin{eqnarray*}
  %y &=& \sum_{k=1}^{K} \delta_{k} \beta_k^T x + \epsilon \\
  \E[ y | x ] &=& M_1^T x \\
  \E[ y^2 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^2 + \E[ \epsilon^2 ] \\
  &=& M_2 \odot x^{\otimes 2} + \sigma^2 \\
  \E[ y^3 | x ] &=& \sum_{k=1}^{K} \pi_{k} (\beta_k^T x)^3
            + 3 \E[\epsilon^2] (\sum_{k=1}^{K} \pi_{k} \beta_k^T) x \\
  &=& M_3 \odot x^{\otimes 3} + 3 \sigma^2 M_1.
\end{eqnarray*}

% Details about low rank regression.
%The symmetric structure of $M_2$ and $M_3$ implies that there are
%$\frac{d (d-1)}{2}$ and $\frac{d (d-1) (d-2)}{6}$ degrees of freedom
%respectively.
However, the moments are further constrained to have
a rank of $K$. To exploit this property, we will use trace norm
regularization with a proximal subgradient descent
algorithm\citationneeded to recover $M_2$,
\begin{align}
  %\hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{i=1}^N ((\xn{i})\tp{2} \odot M_2 - (\yn{i})^2)^2 + \lambda \|M\|_*.
  \hat M_2 &= \arg\min_{M_2} \frac{1}{2 N} \sum_{(x,y) \in D} (x\tp{2} \odot M_2 - y^2)^2 + \lambda \|M\|_*.
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Identifiability from moments}

@@

### 2013-02-12 doc/icml2013/model.tex
minor updates

@@

The mixture of linear regressions model describes response variables $y
\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
Each mode is independently described by regression coefficients
$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
the generative process for the response variables is,
\begin{eqnarray*}
  h &\sim& \mult(\pi) \\
  y | h = k &\sim& \normal{ \beta_{k}^T x}{ \sigma^2 },
\end{eqnarray*}
where $h$ is a latent variable corresponding to the mode.

@@

>>>>

\subsection{Notation}

%We have already defined $y$ to be the response variable and $x \in
%\Re^d$ to be the covariates it depends on. Note that the matrix of
%regressors $B$ is a $d \times K$ matrix, where the $k$-th column is
%$\beta_k$.

We will use the notation $x\tp{p}$ to describe the $p$-th order tensor
formed by taking the outer product of $x$; i.e. $x\tp{p}_{i_1 i_2 \ldots
i_p} = x_{i_1} x_{i_2} \cdots x_{i_p}$. We will use $\odot$ to denote
the general dot product between $p$-th order tensors, i.e. $X \odot
Y = \sum_{i_1, i_2, \ldots i_p} X_{i_1, i_2, \ldots i_p} Y_{i_1, i_2,
\ldots i_p}.$ Similar to how $x^T A$ is the projection of a vector $x$
onto the row space of a matrix $A$, when $X \in \Re^{d^p}$ and $Y \in
\Re^{d^{p+q}}$,  with $p, q > 0$, $X \odot Y \in \Re^{d^q}$ is the
projection of $X$ onto the first $p$ dimensions of $Y$, i.e. $(X \odot
Y)_{i_1, \ldots, i_q} = \sum_{j_1, \ldots, j_p} X_{j_1, \ldots, j_p}
Y_{j_1, \ldots, j_p, i_1, \ldots, i_q}$.

The expected moments of the parameters $\beta_1, \dots, \beta_k$ play an
important role in our algorithm and we will use $M_p = \E[\beta\tp{p}]
= \sum_{k=1}^{K} \pi_k \beta_k^{\otimes p}$ to denote them.

@@

The mixture of linear regressions model describes response variables $y
\in \Re$ dependent on covariates $x \in \Re^d$ through $K$ different
modes which occur with probability $\pi_1, \dots, \pi_K$ respectively.
Each mode is independently described by regression coefficients
$\beta_k$ and we consider a common variance term $\sigma^2$. Given $x$,
the generative process for the response variables is,
\begin{eqnarray*}
  h &\sim& \mult(\pi) \\
  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 },
\end{eqnarray*}
where $h$ is a latent variable corresponding to the mode.

@@

Typically people use EM, but the maximum marginal likelihood is non-convex.

Reference mixture of experts

Acknowledge that mixing proporitions can't depend on $x$
